Let's talk about Thurstone & Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2024-11-30 index.qmd
section:issues; process:write,figures

For instance, @fig-discriminal_process illustrates the discriminal process distributions along a quality trait continuum for two written texts. The figure shows that these processes follow a Normal distribution. Moreover, it depicts differences in the texts' positions along the quality trait continuum, where text B is positioned two units further along the continuum than text A, as indicated by their modal discriminal processes ($S_{B}=2$ and $S_{A}=0$). Finally, it highlights differences in the texts' discriminal dispersions ($\sigma_{B}=1$ and $\sigma_{A}=0.5$), showing that text B exhibits a greater variability in its perceived quality than text A, as reflected by its wider distribution.

@@

Thus, the theory assumes that the observed dichotomous outcome arises from the distribution of the difference between the underlying discriminal processes of the stimuli, known as the *discriminal difference*. Since the individual discriminal processes follow a Normal distribution, their difference also follows a Normal distribution. The mode (or mean) of this distribution, representing the (average) relative separation, is given by the difference between the modal discriminal processes of the stimuli $S_{BA} = S_{B} - S_{A}$. Meanwhile, the dispersion of the distribution, reflecting the variability in the relative separation, is calculated as $\sigma_{BA} = \sqrt{\sigma_{B}^{2} + \sigma_{A}^{2} - \rho\sigma_{B}\sigma_{A}}$. Here, $\sigma_{B}$ and $\sigma_{A}$ are the previously defined discriminal dispersions, while $\rho$ measures the correlation between their discriminal processes. This correlation quantifies how much the judge's perception of the trait in one stimulus influences his perception of the same trait in the other.

@@

For instance, when two texts clearly differ in quality, as depicted in @fig-discriminal_process, higher correlations increase the likelihood that the discriminal difference more distinctly favors text B over text A, as indicated by the shaded gray area under the curve in @fig-correlation. Conversely, it is easy to infer that when the two texts have similar or identical quality levels, higher correlations reduce the likelihood that the discriminal difference distinctly favors one text over the other, as the distribution narrows around zero (not depicted in the figure).

>>>>

For instance, @fig-discriminal_process illustrates the discriminal process distributions along a quality trait continuum for two written texts. The figure shows that these processes follow a Normal distribution. Moreover, it depicts differences in the texts' positions along the quality trait continuum, where text B is positioned further along the continuum than text A, as indicated by their modal discriminal processes ($S_{B}$ and $S_{A}$). Finally, it highlights differences in the texts' discriminal dispersions ($\sigma_{B}$ and $\sigma_{A}$), showing that text B exhibits a greater variability in its perceived quality than text A, as reflected by its wider distribution.

@@

Indeed, the theory assumes that the observed dichotomous outcome arises from the distribution of the difference between the underlying discriminal processes of the stimuli, known as the *discriminal difference*. Since the individual discriminal processes follow a Normal distribution, their difference also follows a Normal distribution. The mode (or mean) of this distribution, representing the (average) relative separation, is given by the difference between the modal discriminal processes of the stimuli $S_{BA} = S_{B} - S_{A}$. Meanwhile, the dispersion of the distribution, reflecting the variability in the relative separation, is calculated as $\sigma_{BA} = \sqrt{\sigma_{B}^{2} + \sigma_{A}^{2} - \rho\sigma_{B}\sigma_{A}}$. Here, $\sigma_{B}$ and $\sigma_{A}$ are the previously defined discriminal dispersions, while $\rho$ measures the correlation between their discriminal processes. This correlation quantifies how much the judge's perception of the trait in one stimulus influences his perception of the same trait in the other.

@@

For instance, when two texts clearly differ in quality, as depicted in @fig-discriminal_process, higher correlations increase the likelihood that the discriminal difference more distinctly favors text B over text A, as indicated by the shaded gray area under the curve in @fig-correlation. Conversely, when the two texts have similar or identical quality levels, higher correlations reduce the likelihood that the discriminal difference distinctly favors one text over the other, as the distribution narrows around zero (not depicted in the figure).
